Out-of-Sample Comparisons of Overfit Models

### 2014-03-28 paper.tex
Minor tweaks to in-sample results

The condition~\label{eq:27} requires that the acceptance region of
$\Lambda$ eventually contains any finite cylindar centered at the
origin. Again,~\eqref{eq:15} implies that statistics like the BIC
will always choose the smaller model for some DGPs, even when the
larger model will give more accurate forecasts. Both models may be
overfit, in that both $K_1/T$ and $K_2/T$ may both be positive in the
limit; the key is that $(K_2 - K_1)/T$ is also positive in the limit.

@@

For the $F$-test we have, as before,
\begin{equation*}
  F = \tfrac{T}{s^2 (K_2 - K_1)} \bh{2T}' M_2' M_2 \bh{2T}.
\end{equation*}
We know that $\sqrt{T}(F - 1)$ is asymptotically normal
\citep{Cal:11c} so the test accepts if
\begin{equation}\label{eq:26}
  \tfrac{T}{s^2 (K_2 - K_1)} \bh{2T}' M_2' M_2 \bh{2T}
  \leq 1 + \delta / \sqrt{T},
\end{equation}
where $\delta$ is chosen to determine the size of the test. In other
words, the test accepts if $M_2 \bh{2T}$ falls in the sphere centered
at the origin with radius $s ((K_2 - K_1) / T)^{1/2} +
O_p(1/\sqrt{T})$.

@@

In the picture, the unknown coefficients satisfy
\begin{equation*}
\theta_2'M_2'M_2\theta_2 > \var(\e_t)^{1/2} ((K_2 - K_1) / T)^{1/2}
\end{equation*}
so the center of the conditional distribution of $\bh{2T}$ given $\E_T
\oosB \leq 0$ is outside the acceptance region of the $F$-test. When
this is the case, the conditional probability that $\bh{2T}$ falls in
the rejection region is less than 1/2, since $M_2 \bh{2T}$ is
uniformly distributed on the cylindar
\begin{equation*}
  (M_2 \bh{2T} - M_2 \theta_2)'(M_2 \bh{2T} - M_2 \theta_2) =
  \theta_2'M_2' M_2 \theta_2 + d
\end{equation*}
for some positive $d$.

@@

Let $(\e_{t+h}, x_{t}) \sim i.i.d.\ N(0,I)$ and let $\theta_1 = 0$.
For any $d \geq 0$, the event $\E_T \oosB = -d$ implies that
\begin{equation*}
  \E_T (y_{T+\h+1} - x_{1T+1}'\bh{1T})^2
  = \E_T (y_{T+\h+1} - x_{2T+1}'\bh{2T})^2 - d \quad a.s.
\end{equation*}
which can be expressed as
\begin{equation}\label{eq:34}
  d + \bh{1T}'\bh{1T} - \bh{2T}'M_1'M_1\bh{2T}
  = (\bh{2T} - \theta_2)' M_2'M_2 (\bh{2T} - \theta_2) \quad a.s.
\end{equation}
$M_2 \bh{2T}$ is normally distributed conditional on $\bh{1T}$ and
$M_1 \bh{2T}$, and is distributed on the surface of the sphere defined
by~\eqref{eq:34} conditional on $\bh{1T}$, $M_1 \bh{2T}$, and the
event $\E_T \oosB = -d$.

@@

Now let $\delta > 0$ be an arbitrary but small constant, and choose
$\theta_2$ far enough from the origin to ensure that
\begin{equation*}
  \Pr[\theta_2'M_2' V_T M_2 \theta_2 \leq c] \leq \delta
\end{equation*}
for large enough $T$. Then
\begin{align*}
  \E(\Lambda \mid \E_T \oosB \leq 0) &
  \geq \Pr[\bh{2T}' M_2' V_T M_2 \bh{2T} > c \mid \E_T \oosB \leq 0] \\
  &= \Pr[\bh{2T}' M_2' V_T M_2 \bh{2T} > c \text{~~and~~}
  \theta_2'M_2' V_T M_2 \theta_2 > c \mid \E_T \oosB  \leq 0] \\
  & \quad + \Pr[\bh{2T}' V_T M_2 \bh{2T} > c \text{~~and~~ }
  \theta_2'M_2' V_T M_2 \theta_2 \leq c \mid \E_T \oosB \leq 0].
\end{align*}

The second term is nonnegative by design. By conditioning on
$\theta_2'M_2' V_T M_2 \theta_2 > c$ our earlier argument shows that
the first term is greater than or equal to $(1 - \delta) / 2$. Since
$\delta$ is arbitrarily small, this completes the proof.
\qed

\subsection*{Proof of Theorem~\ref{res:insample2}}
As in the proof of Theorem~\ref{res:insample1}, let $(\e_{t+h}, x_{t})
\sim i.i.d.\ N(0,I)$ and let $\theta_1 = 0$. The event $\E_T \oosB
\geq 0$ implies that
\begin{equation*}
  \E_T (y_{T+\h+1} - x_{1T+1}'\bh{1T})^2
  \geq \E_T (y_{T+\h+1} - x_{2T+1}'\bh{2T})^2 \quad a.s.
\end{equation*}
which can be expressed as
\begin{equation}\label{eq:34}
  \bh{1T}'\bh{1T} - \bh{2T}'M_1'M_1\bh{2T}
  \geq (\bh{2T} - \theta_2)' M_2'M_2 (\bh{2T} - \theta_2) \quad a.s.
\end{equation}
$M_2 \bh{2T}$ is normally distributed conditional on $\bh{1T}$ and
$M_1 \bh{2T}$, and is a.s. contained in the sphere defined
by~\eqref{eq:34} conditional on $\bh{1T}$, $M_1 \bh{2T}$, and the
event $\E_T \oosB \geq 0$.

Let $\delta$ be a small arbitrary constant, define $\theta^*$ as
\begin{gather*}
  \theta^* = \argmax_\theta \theta' M_2' M_2 \theta \\
  s.t.\ \bh{1T}'\bh{1T} - \bh{2T}'M_1'M_1\bh{2T} = (\theta - \theta_2)' M_2'M_2 (\theta - \theta_2)
\end{gather*}
($\theta^*$ depends on $\bh{1T}$ and $M_1 \bh{2T}$) and choose $c$ so
that
\begin{equation*}
  \Pr[ \theta^{* \prime} M_2 M_2 \theta^* > c] < \delta.
\end{equation*}

>>>>

The condition~\ref{eq:27} requires that the acceptance region of
$\Lambda$ eventually contains any finite cylindar centered at the
origin. Again,~\eqref{eq:15} implies that statistics like the BIC
will always choose the smaller model for some DGPs, even when the
larger model will give more accurate forecasts. Both models may be
overfit, in that both $K_1/T$ and $K_2/T$ may both be positive in the
limit; the key is that $(K_2 - K_1)/T$ is also positive in the limit.

@@

For the $F$-test we have, as before,
\begin{equation*}
  F = \tfrac{T}{s^2 (K_2 - K_1)} \bh{2T}' M_2' M_2 \bh{2T}.
\end{equation*}
We know that $\sqrt{T}(F - 1)$ is asymptotically normal since its
numerator is Chi-squared with $K_2-K_1$ degrees of freedom and obeys a
CLT.  The test accepts if
\begin{equation}\label{eq:26}
  \tfrac{T}{s^2 (K_2 - K_1)} \bh{2T}' M_2' M_2 \bh{2T}
  \leq 1 + \delta / \sqrt{T},
\end{equation}
where $\delta$ is chosen to determine the size of the test. In other
words, the test accepts if $M_2 \bh{2T}$ falls in the sphere centered
at the origin with radius $s ((K_2 - K_1) / T)^{1/2} +
O_p(1/\sqrt{T})$.

@@

In the picture, the true coefficients $\theta_2$ satisfy
\begin{equation*}
\theta_2'M_2'M_2\theta_2 > \var(\e_t)^{1/2} ((K_2 - K_1) / T)^{1/2}
\end{equation*}
so the center of the conditional distribution of $\bh{2T}$ given $\E_T
\oosB \leq 0$ is outside the acceptance region of the $F$-test. When
this is the case, the conditional probability that $\bh{2T}$ falls in
the rejection region is less than 1/2, since $M_2 \bh{2T}$ is
uniformly distributed on the cylindar
\begin{equation*}
  (M_2 \bh{2T} - M_2 \theta_2)'(M_2 \bh{2T} - M_2 \theta_2) =
  \theta_2'M_2' M_2 \theta_2 + d
\end{equation*}
for some positive $d$.

@@

Let $(\e_{t+h}, x_{t}) \sim i.i.d.\ N(0,I)$ and let $\theta_1 = 0$.
For any $d \geq 0$, the event $\E_T \oosB = -d$ implies that
\begin{equation*}
  \E_T (y_{T+\h+1} - x_{1T+1}'\bh{1T})^2
  = \E_T (y_{T+\h+1} - x_{2T+1}'\bh{2T})^2 - d \quad a.s.
\end{equation*}
which can be expressed as
\begin{equation}\label{eq:34}
  d + \theta_2'M_2' M_2 \theta_2 + \bh{1T}'\bh{1T} - \bh{2T}'M_1'M_1\bh{2T}
  = (\bh{2T} - \theta_2)' M_2'M_2 (\bh{2T} - \theta_2) \quad a.s.
\end{equation}
$M_2 \bh{2T}$ is normally distributed conditional on $\bh{1T}$ and
$M_1 \bh{2T}$, and is distributed on the surface of the sphere defined
by~\eqref{eq:34} conditional on $\bh{1T}$, $M_1 \bh{2T}$, and the
event $\E_T \oosB = -d$.

@@

Now let $\delta > 0$ be an arbitrary but small constant, and choose
$\theta_2$ far enough from the origin to ensure that
\begin{equation*}
  \Pr[\theta_2'M_2' V_T M_2 \theta_2 \leq c] \leq \delta
\end{equation*}
for large enough $T$. Then
\begin{align*}
  \E(\Lambda \mid \E_T \oosB \leq 0) &
  \geq \Pr[\bh{2T}' M_2' V_T M_2 \bh{2T} > c \mid \E_T \oosB \leq 0] \\
  &= \Pr[\bh{2T}' M_2' V_T M_2 \bh{2T} > c \text{~~and~~}
  \theta_2'M_2' V_T M_2 \theta_2 > c \mid \E_T \oosB  \leq 0] \\
  & \quad + \Pr[\bh{2T}' V_T M_2 \bh{2T} > c \text{~~and~~ }
  \theta_2'M_2' V_T M_2 \theta_2 \leq c \mid \E_T \oosB \leq 0].
\end{align*}
The second term is nonnegative by design. By conditioning on
$\theta_2'M_2' V_T M_2 \theta_2 > c$ our earlier argument shows that
the first term is greater than or equal to $(1 - \delta) / 2$. Since
$\delta$ is arbitrarily small, this completes the proof.
\qed

\subsection*{Proof of Theorem~\ref{res:insample2}}
As in the proof of Theorem~\ref{res:insample1}, let $(\e_{t+h}, x_{t})
\sim i.i.d.\ N(0,I)$ and let $\theta_1 = 0$. The event $\E_T \oosB
\geq 0$ implies that
\begin{equation*}
  \E_T (y_{T+\h+1} - x_{1T+1}'\bh{1T})^2
  \geq \E_T (y_{T+\h+1} - x_{2T+1}'\bh{2T})^2 \quad a.s.
\end{equation*}
which can be expressed as
\begin{equation}\label{eq:2}
  \theta_2'M_2 M_2'\theta_2 + \bh{1T}'\bh{1T} - \bh{2T}'M_1'M_1\bh{2T}
  \geq (\bh{2T} - \theta_2)' M_2'M_2 (\bh{2T} - \theta_2) \quad a.s.
\end{equation}
$M_2 \bh{2T}$ is normally distributed conditional on $\bh{1T}$ and
$M_1 \bh{2T}$, and is a.s. contained in the sphere defined
by~\eqref{eq:2} conditional on $\bh{1T}$, $M_1 \bh{2T}$, and the
event $\E_T \oosB \geq 0$.

Let $\delta$ be a small arbitrary constant, define $\theta^*$ as
\begin{gather*}
  \theta^* = \argmax_\theta \theta' M_2' M_2 \theta \\
  s.t.\quad \theta_2'M_2'M_2\theta_2 +  \bh{1T}'\bh{1T} - \bh{2T}'M_1'M_1\bh{2T} = (\theta - \theta_2)' M_2'M_2 (\theta - \theta_2)
\end{gather*}
($\theta^*$ is stochastic and depends on $\bh{1T}$ and $M_1 \bh{2T}$)
and choose $c$ so that
\begin{equation*}
  \Pr[ \theta^{* \prime} M_2 M_2 \theta^* > c] < \delta.
\end{equation*}
